% !TEX root= ../main.tex
\section{Odd paths}
\label{sec:Odd paths}
Given two vertices $x_1$ and $x_2$ in a graph $G$, if we in Neg are able to prove $\ol{x_1x_2}$ from the axioms we get from $G$, we say that $x_1$ and $x_2$ are \textit{vel-connected}.
This section will try to answer the question ``what kind of graph structures imply that certain vertices are vel-connected?''.

Because of soundness of Neg, if two vertices are vel-connected in a graph, for any kernel $K$ in that graph, at least one of the two vertices are outside $K$.
The counterpositive of this observation being that for a graph $G$, if there exists a kernel $K \subseteq G$ such that two vertices, $x_1$ and $x_2$ are both in $K$, then $x_1$ and $x_2$ cannot be vel-connected in $G$.
This fact will be used when arguing why some graph structures are \textit{not} giving us vel-connected vertices.

We already know that whenever two vertices are connected by an edge, their binary NAND is trivially provable in Neg, since it is a part of the axioms.

We illustrate this case with the figure below, where dashed lines represent possible out-edges to irrelevant parts of the graph.
If a vertex has no dashed edges, it means that we disallow any additional edges out from this vertex.
\[
  \begin{tikzpicture}
    [
    point/.style={circle,draw,inner sep=0pt,minimum size=2mm},
    ]
    \node (1) at (1,1) [point] {};
    \node (2) at (2,1) [point] {};
    \draw [-latex] (1) to (2);

    \node (e1) [above left=4mm and 6mm of 1]  {};
    \node (e2) [left=8mm of 1] {};
    \node (e3) [below left=4mm and 6mm of 1] {};
    \node (e4) [above right=4mm and 6mm of 2] {};
    \node (e5) [right=8mm of 2] {};
    \node (e6) [below right=4mm and 6mm of 2] {};
    \draw [dashed] (e1) to (1);
    \draw [dashed] (e2) to (1);
    \draw [dashed] (e3) to (1);
    \draw [dashed] (e4) to (2);
    \draw [dashed] (e5) to (2);
    \draw [dashed] (e6) to (2);
  \end{tikzpicture}
\]
The basic example above tells us that two vertices are vel-connected when one is the successor of the other.
It is however easy to find a case examplifying how the implication does not hold in the opposite direction.
\[
  \begin{tikzpicture}
    [
    point/.style={circle,draw,inner sep=0pt,minimum size=2mm},
    ]
    \node (1) at (0,1) [point,label=above:$x_0$] {};
    \node (2) at (1,1) [point,label=above:$x_1$] {};
    \node (3) at (2,1) [point,label=above:$x_2$] {};
    \node (4) at (3,1) [point,label=above:$x_3$] {};
    \draw [-latex] (1) to (2);
    \draw [-latex] (2) to (3);
    \draw [-latex] (3) to (4);

    \node (e4) [above right=4mm and 6mm of 4] {};
    \node (e5) [right=8mm of 4] {};
    \node (e6) [below right=4mm and 6mm of 4] {};
    \draw [dashed] (e4) to (4);
    \draw [dashed] (e5) to (4);
    \draw [dashed] (e6) to (4);
  \end{tikzpicture}
\]
The above graph provides us the following axioms:
$NAND = \{\ol{x_0x_1},\ol{x_1x_2},\ol{x_2x_3}\}$, $OR = \{x_0x_1,x_1x_2,x_2x_3\}$.
From these axioms, we can now, despite the fact that the vertices $x_0$ and $x_3$ are not connected by an edge, easily prove the NAND-clause $\ol{x_0x_3}$:

\begin{prooftree*}
  \Hypo{\ol{x_0x_1}}
  \Hypo{\ol{x_2x_3}}
  \Infer[left label=$x_1x_2$]2{\ol{x_0x_3}}
\end{prooftree*}

Intuitively, one can imagine that the proof above is connecting two NAND-clauses using an OR-clause, resulting in a new binary NAND-clause containing vertices that are weaklier connected than the ones we started with.
This can be done repeatedly, resulting in the ability to prove binary NAND-clauses from vertices that are connected by arbitrarily long paths of the kind above.

Consider the following graph:
\[
  \begin{tikzpicture}
    [
    point/.style={circle,draw,inner sep=0pt,minimum size=2mm},
    ]
    \node (1) at (0,1) [point,label=above:$x_0$] {};
    \node (2) at (1,1) [point,label=above:$x_1$] {};
    \node (3) at (2,1) [point,label=above:$x_2$] {};
    \node (4) at (3,1) [point,label=above:$x_3$] {};
    \node (5) at (4,1) [point,label=above:$x_4$] {};
    \node (6) at (5,1) [point,label=above:$x_5$] {};
    \node (7) at (6,1) [point,label=above:$x_6$] {};
    \node (8) at (7,1) [point,label=above:$x_7$] {};
    \node (9) at (8,1) [point,label=above:$x_8$] {};
    \node (10) at (9,1) [point,label=above:$x_9$] {};
    \draw [-latex] (1) to (2);
    \draw [-latex] (2) to (3);
    \draw [-latex] (3) to (4);
    \draw [-latex] (4) to (5);
    \draw [-latex] (5) to (6);
    \draw [-latex] (6) to (7);
    \draw [-latex] (7) to (8);
    \draw [-latex] (8) to (9);
    \draw [-latex] (9) to (10);

    \node (e4) [above right=4mm and 6mm of 10] {};
    \node (e5) [right=8mm of 10] {};
    \node (e6) [below right=4mm and 6mm of 10] {};
    \draw [dashed] (e4) to (10);
    \draw [dashed] (e5) to (10);
    \draw [dashed] (e6) to (10);
  \end{tikzpicture}
\]

With axioms from the above graph, $\ol{x_0x_9}$ can be proven in Neg in the following way:

\begin{prooftree*}
  \Hypo{\ol{x_0x_1}}
  \Hypo{\ol{x_2x_3}}
  \Infer[left label=$x_1x_2$]2{\ol{x_0x_3}}
  \Hypo{\ol{x_4x_5}}
  \Infer[left label=$x_4x_5$]2{\ol{x_0x_5}}
  \Hypo{\ol{x_6x_7}}
  \Infer[left label=$x_6x_7$]2{\ol{x_0x_7}}
  \Hypo{\ol{x_8x_9}}
  \Infer[left label=$x_8x_9$]2{\ol{x_0x_9}}
\end{prooftree*}

Observe that the above proof is also proving $\ol{x_0x_3}$, $\ol{x_0x_5}$ and $\ol{x_0x_7}$ along the way, all of which contain vertices connected by paths of \textit{odd} length.
This is an important point.
NAND-clauses containing vertices connected only by paths of \textit{even} length cannot be proven in the same manner as above.

In many cases, such NAND-clauses cannot be proven at all.
This is examplified by the below graph, with a kernel containing the vertices $a$ and $b$, showing that the NAND $\ol{ab}$ is unprovable in Neg.
The kernel in the graph below is represented by the black vertices.

\[
  \begin{tikzpicture}
    [
    point/.style={circle,draw,inner sep=0pt,minimum size=2mm},
    one/.style={fill=black},
    ]
    \node (1) at (0,1) [point, one, label=above:$a$] {};
    \node (2) at (1,1) [point] {};
    \node (3) at (2,1) [point, one] {};
    \node (4) at (3,1) [point] {};
    \node (5) at (4,1) [point, one, label=above:$b$] {};
    \draw [-latex] (1) to (2);
    \draw [-latex] (2) to (3);
    \draw [-latex] (3) to (4);
    \draw [-latex] (4) to (5);

    \node (e4) [above right=4mm and 6mm of 5] {};
    \node (e5) [right=8mm of 5] {};
    \node (e6) [below right=4mm and 6mm of 5] {};
    \draw [dashed] (e4) to (5);
    \draw [dashed] (e5) to (5);
    \draw [dashed] (e6) to (5);
  \end{tikzpicture}
\]

Restricting our paths to be of odd length lets us avoid cases like the one above, but there are still some cases left to avoid, in order for our implication to hold.

All the paths so far have been free for any branching, i.e. all vertices in the paths have been of degree 1.
We will call such paths \textit{fully trimmed paths}, and formally define them as paths where for each pair of vertices $x$ and $y$ consecutive in the path, $N(x) = \{y\}$.

If two vertices $a$ and $b$ are connected by an odd path that is \textit{not} fully trimmed, $\ol{ab}$ is not neseccarily provable in Neg.
The below graph examplifies this with a kernel containing both $a$ and $b$.

\[
  \begin{tikzpicture}
    [
    point/.style={circle,draw,inner sep=0pt,minimum size=2mm},
    one/.style={fill=black},
    ]
    \node (1) at (0,1) [point, one, label=above:$a$] {};
    \node (2) at (1,1) [point] {};
    \node (2b) [point, one, below right=4mm and 6mm of 2] {};
    \node (3) at (2,1) [point] {};
    \node (4) at (3,1) [point, one,label=above:$b$] {};
    \draw [-latex] (1) to (2);
    \draw [-latex] (2) to (3);
    \draw [-latex] (2) to (2b);
    \draw [-latex] (3) to (4);

    \node (e4) [above right=4mm and 6mm of 4] {};
    \node (e5) [right=8mm of 4] {};
    \node (e6) [below right=4mm and 6mm of 4] {};
    \draw [dashed] (e4) to (4);
    \draw [dashed] (e5) to (4);
    \draw [dashed] (e6) to (4);
  \end{tikzpicture}
\]

Generalizing this, we get that whenever two vertices, $x_0$ and $x_k$, are connected by a fully trimmed path of odd length, they are vel-connected.
Their corresponding NAND-clause $\ol{x_0x_k}$ can be proven in Neg in the following way:
\begin{prooftree*}
  \Hypo{\ol{x_0x_1}}
  \Hypo{\ol{x_2x_3}}
  \Infer[left label=$x_1x_2$]2{\ol{x_0x_3}}
  \Hypo{\ol{x_4x_5}}
  \Infer[left label=$x_3x_4$]2{\ol{x_0x_5}}
  \Ellipsis{}{\ol{x_0x_{k-2}}}
  \Hypo{\ol{x_{k-1}x_k}}
  \Infer[left label=$x_{k-2}x_{k-1}$]2{\ol{x_0x_k}}
\end{prooftree*}

All the OR-clauses used are indeed binary, letting us prove our NAND-clause without introducing any other clauses than the ones we get from the path itself.
However, notice that only half of the OR-clauses is actually in use in such a proof.
Thus, we need only to restrict half of the vertices in the path to not branch.

With the above proof in mind, consider the following graph:
\[
  \begin{tikzpicture}
    [
    point/.style={circle,draw,inner sep=0pt,minimum size=2mm},
    ]
    \node (1) at (0,1) [point,label=above:$x_0$] {};
    \node (2) at (1,1) [point,label=above:$x_1$] {};
    \node (3) at (2,1) [point,label=above:$x_2$] {};
    \node (4) at (3,1) [point,label=above:$x_3$] {};
    \node (dots) at (4,1) [] {\dots};
    \node (k2) at (5,1) [point,label=above:$x_{k-2}$] {};
    \node (k1) at (6,1) [point,label=above:$x_{k-1}$] {};
    \node (k) at (7,1) [point,label=above:$x_k$] {};
    \draw [-latex] (1) to (2);
    \draw [-latex] (2) to (3);
    \draw [-latex] (3) to (4);
    \draw [-latex] (4) to (dots);
    \draw [-latex] (dots) to (k2);
    \draw [-latex] (k2) to (k1);
    \draw [-latex] (k1) to (k);

    \node (e4) [above right=4mm and 6mm of k] {};
    \node (e5) [right=8mm of k] {};
    \node (e6) [below right=4mm and 6mm of k] {};
    \draw [dashed] (e4) to (k);
    \draw [dashed] (e5) to (k);
    \draw [dashed] (e6) to (k);

    \node (b1) [below right=4mm and 6mm of 1] {};
    \node (b3) [below right=4mm and 6mm of 3] {};
    \node (bk1) [below right=4mm and 6mm of k1] {};
    \draw [dashed] (1) to (b1);
    \draw [dashed] (3) to (b3);
    \draw [dashed] (k1) to (bk1);
  \end{tikzpicture}
\]
In the above graph, only every other vertex are restricted to only have one successor.
We will call this path variant an \textit{oddly trimmed path}, and define it formally as follows:
A path $\langle x_0,x_k\rangle$\todo{This notation is not explained} is oddly trimmed if for each odd $i < k$, $N(x_i) = \{x_{i+1}\}$.
One can immediately note that any fully trimmed path is also oddly trimmed.

The axioms we get from the oddly trimmed path above does not differ significantly from the fully trimmed variant.
Since the vertices $x_0, x_2, x_4, \dots ,x_{k_1}$ no longer have single successors, their corresponding OR-clauses will no longer be binary.
However, since none of these OR-clauses are used in the above proof, the proof will remain valid also for the oddly trimmed path.

We can thus generalize further and say that any two vertices connected by an oddly trimmed path of odd length is vel-connected.
