% !TEX root= ../main.tex
\subsection{Discourse Theories and Digraphs}
\label{sub:Discourse Theories and Digraphs}
As mentioned earlier, there is a close connection between (1) models of a discourse, (2) kernels of a graph and (3) solutions of a graph.
While we have the equivalence between (2) and (3), we will now look at two functions connecting (1) and (2).
This correspondence was shown by Roy T. Cook in \cite{cook}.
We get the following definitions from \cite{apal-digraph}

$\mathcal{T}:$ translating a digraph \textbf{G} into a corresponding theory $\mathcal{T}(\mathbf{G})$ such that $sol(\mathbf{G}) = mod(\mathcal{T}(G))$.

$\mathcal{G}:$ translating a theory $T$ into a corresponding digraph $\mathcal{G}(T)$ such that $mod(T) = sol(\mathcal{G}(T))$.

Given any digraph \textbf{G} we get the theory $\mathcal{T}(\mathbf{G})$ by taking, for each $x \in G$, the formula $x \leftrightarrow \bigwedge_{y \in N(x)} \neg y$ where $\bigwedge \emptyset = 1$.\\

\begin{example}\label{ex:3graphs}
  \[
    \begin{tikzpicture}
      [
      point/.style={circle,draw,inner sep=0pt,minimum size=2mm},
      collection/.style={thick,rectangle,draw,inner sep=0pt,minimum height=14mm, minimum width= 9mm}
      ]
      \node (label) at (0,1) [label=left:$(\mathbf{G_1})\;$] {};
      \node (0) at (0,1) [point,label=above:a] {};
      \draw [-latex, loop right] (0) to (0);

      \node (label) at (3,1) [label=left:$(\mathbf{G_2})\;$] {};
      \node (0) at (4,1) [point,label=above:a] {};
      \node (1) at (3,1) [point,label=above:b] {};
      \node (2) at (5,1) [point,label=above:c] {};
      \draw [-latex] (0) to (1);
      \draw [-latex] (0) to (2);

      \node (label) at (8,1) [label=left:$(\mathbf{G_3})\;$] {};
      \node (0) at (8,1) [point,label=above:a] {};
      \node (1) at (8,0) [point,label=left:b] {};
      \node (2) at (9,0) [point,label=right:c] {};
      \draw [-latex] (0) to (2);
      \draw [-latex] (0) to (1);
      \draw [-latex, bend right] (1) to (2);
      \draw [-latex, bend right] (2) to (1);
    \end{tikzpicture}
  \]
  Using the graphs from above, we get the following theories using $\mathcal{T}$:
  \begin{align}
    \mathcal{T}(\mathbf{G_1}) &= \big \{ a \leftrightarrow \neg a \big \} \\
    \mathcal{T}(\mathbf{G_2}) &= \big \{ a \leftrightarrow (\neg b \wedge \neg c), b, c \big \}\\
    \mathcal{T}(\mathbf{G_3}) &= \big \{ a \leftrightarrow (\neg b \wedge \neg c), b \leftrightarrow \neg c, c \leftrightarrow \neg b \big \}
  \end{align}

\end{example}
The fact that $sol(\mathbf{G}) = mod(\mathcal{T}(G))$ is shown in \cite{apal-digraph}.
Allthough not proving it, we can observe that $\mathbf{G_1}$ has no solution, just like its corresponding theory $\mathcal{T}(\mathbf{G_1})$ has no satisfying models.
$\mathbf{G_2}$ has one solution, where one assigns $a=0, b=1, c=1$.
This assignment also works as a satisfying model for $\mathcal{T}(\mathbf{G_2})$.
In $\mathbf{G_3}$, we get two solutions, both with $a$ assigned to 0, but with 0 and 1 distributed on $b$ and $c$.
These are also the only two models satisfying $\mathcal{T}(\mathbf{G_3})$.

Conversely, given any discourse theory $T$ (in fact, this will work given any PL theory, since we can translate CNF to GNF), we can derive the corresponding graph $\mathcal{G}(T)$ in the following way:
All variables in the theory are vertices, and for each formula $x \leftrightarrow \bigwedge_{i \in I_x} y_i$ make a directed edge $\langle x,y_i \rangle$ for each $i \in I_x$.\\

\begin{example}
  \begin{align}
    T_1 &= \neg a &&\iff ( a \lar \neg a'), (a' \lar \neg a), (y_1 \lar (\neg a' \wedge \neg y_1))\\
    T_2 &= a \vee b &&\iff ( a \lar \neg a'), (a' \lar \neg a), (b \lar \neg b'), (b' \lar \neg b), (y_1 \lar (\neg a \wedge \neg b \wedge \neg y_1))
  \end{align}
  Using $\mathcal{G}$ on the above theories -- translated to GNF -- gives us the following graphs:
  \[
    \begin{tikzpicture}
      [
      point/.style={circle,draw,inner sep=0pt,minimum size=2mm},
      collection/.style={thick,rectangle,draw,inner sep=0pt,minimum height=14mm, minimum width= 9mm}
      ]
      \node (label) at (0,1) [label=left:$(\mathcal{G}(T_1))\;$] {};
      \node (0) at (0,1) [point,label=above:$y_1$] {};
      \node (1) at (1,1) [point,label=above:a'] {};
      \node (2) at (2,1) [point,label=above:a] {};
      \draw [-latex, loop below] (0) to (0);
      \draw [-latex] (0) to (1);
      \draw [-latex, bend right] (1) to (2);
      \draw [-latex, bend right] (2) to (1);

      \node (label) at (5,1) [label=left:$(\mathcal{G}(T_1))\;$] {};
      \node (0) at (5,1) [point,label=above:$y_1$] {};
      \node (1) at (6,0) [point,label=above:b] {};
      \node (2) at (7,0) [point,label=above:b'] {};
      \node (3) at (6,2) [point,label=above:a] {};
      \node (4) at (7,2) [point,label=above:a'] {};
      \draw [-latex, loop below] (0) to (0);
      \draw [-latex] (0) to (1);
      \draw [-latex, bend right] (1) to (2);
      \draw [-latex, bend right] (2) to (1);
      \draw [-latex] (0) to (3);
      \draw [-latex, bend right] (3) to (4);
      \draw [-latex, bend right] (4) to (3);
    \end{tikzpicture}
  \]
\end{example}
Again, will we not be proving the correspondence, but notice that $T_1$ has one model satisfying it, where $a = 0$.
$\mathcal{G}(T_1)$ also has one solution, namely where $a = 0, a' = 1, y_1 = 0$.
$T_2$ has three solutions, where either $a$, $b$ or both are assigned to $1$.
This reflects onto the graph since $y_1$ has to be assigned to $0$, thus forcing $a$ or $b$ to be assigned to $1$.
The fact that $\mathcal{G}$ gives us the correspondence we are looking for is shown in \cite{apal-digraph}.

With the problem of solutions in the graph being equivalent with SAT, we get our final equivalence between kernels in the graph and satisfying models of the theory.
This equivalence connects the fields of logic with the fields of graph theory:
A graph has a kernel if and only if its corresponding discourse is consistent (non-paradoxical).
A theory is paradoxical if and only if its corresponding graph has no kernels.
Because of this tight link, we will often refer to graphs without kernels as paradoxical graphs.

The applicability of kernels should by now be obvious.
In the next section we will explore some of the various findings within Kernel Theory, and especially the findings related to infinitary graphs.
