% !TEX root= ../main.tex
\section{Graphs, Kernels and Solutions}
\label{sec:Graphs, Kernels and Solutions}
\begin{definition}
  A \textit{directed graph (digraph)} is a pair \textbf{G} = $\langle G,N \rangle$ where $G$ is a set of vertices while $N \subseteq G \times G$ is a binary relation representing the edges in \textbf{G}.
\end{definition}
We say that the graph is \textit{finite} if its set of vertices is finite; \textit{infinite} otherwise.

We use the notation $N(x)$ to denote the set of all vertices that are targeted by edges originating in $x$ (\textit{successors} of $x$).
Similarly, $N^-(x)$ denotes the set of all vertices with edges targeting $x$ (\textit{predecessors} of $x$).
We define the two functions formally as follows:
\begin{align}
  N(x) := \{y \;|\; (x,y) \in N\}\\
  N^-(x) := \{ y \;|\; (y,x) \in N \}
\end{align}
\begin{definition}
  The number of successors a vertex has is often called the \textit{out-degree} of that vertex.
  The number of predecessors is called the \textit{in-degree}.
  If a vertex has out-degree 0, we call it a \textit{sink}; if it has in-degree 0, we call it a \textit{source}.
  If the out-degree of each vertex in a graph is finite, we say that the graph is \textit{finitary}.
  A graph that is not finitary is called \textit{infinitary}.
\end{definition}
Note the difference between the concept of a finite graph and the one of a finitary graph;
an infinitary graph is infinite, but a finitary graph is not necessarily finite.
\begin{definition}
  A \textit{simple path} is a sequence of distinct vertices $x_1,x_2,\dots,x_n$ such that for each consecutive pair $x_i,x_{i+1}$ from the sequence, we have $(x_i, x_{i+1}) \in N$.
  We say that two paths are \textit{disjoint} if they do not share any vertices (with the possible exception of their initial vertices).
\end{definition}
The number of edges in the path is called the \textit{length} of the path.

The functions $N$ and $N^-$ can be extended pointwise to sets in the following way:
\begin{align}
  N(X) = \bigcup_{x \in X} N(x)\\
  N^-(X) = \bigcup_{x \in X} N(x)
\end{align}

\begin{definition}
  A \textit{kernel} is a set of vertices $K \subseteq G$ such that:
  \begin{align}
    G \setminus K = N^-(K)
  \end{align}
\end{definition}
The equivalence in the above definition can be split into two inclusions to be more easily understood:

$G \setminus K \subseteq N^-(K)$, saying that each vertex outside the kernel has an edge into the kernel (K is \textit{absorbing}).
A consequence of this is that a kernel has to be non-empty, unless the graph is empty.

$N^-(K) \subseteq G \setminus K$, saying that each edge targeting a vertex within the kernel has to come from outside, thus no two vertices in the kernel are connected by an edge (K is \textit{independent}).

Kernels have been studied over several decades, not only in graph theory, but also within the fields of game theory and economics.
The concept was first defined and used by von Neumann and Morgenstern in \cite{neumann}.

In a graph representing some sort of a turn-based game, where vertices are states and edges are transitions between states, one can often work out winning strategies whenever one finds a kernel in the graph.
Whenever one is outside of the kernel, one always has the possibility of moving inside the kernel (since the kernel is absorbing), while inside the kernel one \textit{has} to move out of it (since the kernel is independent).
If you are the player with the choice outside the kernel, you can control the game and choose to stabilize it by always moving into the kernel, forcing the opponent to move out again on the next turn.

Deciding whether kernels exist in finite graphs has been shown to be an NP-complete problem\cite{chvatal}.
This should not be surprising, since we are in the middle of showing the equivalence between this problem and the problem of finding satisfying models of PL-theories (SAT), which we know is NP-complete.
It should be noted however, that in this thesis, we will mainly concern ourselves with the existence of kernels in \textit{infinitary} graphs and thus also SAT over infinitary formulae.

We will get the correspondence between models of a discourse theory and kernels in a graph through an alternative, equivalent kernel definition called a \textit{solution}.
\begin{definition}
  Given a directed graph \textbf{G} = $\langle G,N \rangle$, an \textit{assignment} $\alpha \in 2^G$ is a function mapping every vertex in the graph to either 0 or 1.
  A \textit{solution} is an assignment $\alpha$ such that for all $x \in G:$
  \begin{align}
    \alpha(x) = 1 \iff \alpha(N(x)) = \{ 0 \}
  \end{align}
\end{definition}
This means that for any node $x$, if $x$ is assigned 1, then all its successors have to be assigned  0, and if $x$ is assigned 0, then there has to exist a node assigned 1 among its successors.
A consequence of this definition is that all sink nodes (nodes with no outgoing edges) in the graph have to be assigned 1, since it vacuously does not point to any node assigned 1.
We use the notation $sol(\mathbf{G})$ to denote the set of all solutions of the graph \textbf{G}.

We get the equivalence between kernels and solutions from the following two facts:
Given a solution, the set of all vertices assigned 1 is a kernel.
Given a kernel, the function assigning 1 to all vertices in the kernel and 0 to the rest, is a solution.
